---
title: "localmiic"
author: "achemkhi"
date: "2025-02-02"
output: html_document



---

# local miic 

- miic_run_analysis_test.R command line 
> Rscript /home/miic/miic_webserver/miic_run_analysis_test.R 
    -i /home/miic/server/server_outputs/Y7qeg1ZJMyebZgQ2eeT1/input.tsv 
    -o /home/miic/server/server_outputs/Y7qeg1ZJMyebZgQ2eeT1 
    -l orientation 
    -e -1 
    --complexity nml 
    --conf_shuffle 0 
    --conf_ratio 0 
    --consistency no 
    --orient_probab_ratio 1.0
    --n_threads 3 
    --plots 
    --orientation 
    --test_MAR -s /home/miic/server/server_outputs/Y7qeg1ZJMyebZgQ2eeT1/clean_state_order.tsv 
    --miic_mode S 
    --max_nodes 200


# command line to compile the package
```{bash}
cd /Users/alichemkhi/Desktop/code/miic_R_package;
R CMD INSTALL . --preclean --library="/Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/myLibraries"
```

## toy dataset

```{r}
library(miic, lib.loc="/Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/myLibraries")
library(sessioninfo)
#package_info("miic")

input_file="/Users/alichemkhi/Desktop/myProjects/miic_helper/data/ibreg_PC_smart_selection.9.csv"
input_st="/Users/alichemkhi/Desktop/myProjects/miic_helper/data/ibreg_PC_smart_selection.9.csv"

df <- read.table(input_file, sep = ",", header = T,
                 stringsAsFactors = FALSE, check.names = FALSE)[1:1000,]
state.order <- read.table(input_st, sep = ",", header = T,
                 stringsAsFactors = FALSE, check.names = FALSE)

miic_res <- miic(df,
                 state_order=state.order,
                 orientation=T,
                 n_threads=6,
                 verbose=FALSE)


```

## segfault bug 

```{r }
library(miic, lib.loc="/Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/myLibraries")
library(sessioninfo)
package_info("miic")

input_file="/Users/alichemkhi/Desktop/myProjects/iBregs_induction/outputs/miic_outputs/ibreg_PC/old_selection/ibreg_PC.133.csv"
input_st="/Users/alichemkhi/Desktop/myProjects/iBregs_induction/outputs/miic_outputs/ibreg_PC/old_selection/ibreg_PC.133.st.txt"

df <- read.table(input_file, sep = "\t", header = T,
                 stringsAsFactors = FALSE, check.names = FALSE)
state.order <- read.table(input_st, sep = "\t", header = T,
                 stringsAsFactors = FALSE, check.names = FALSE)


miic_res <- miic(df,
                 state_order=state.order,
                 orientation=T,
                 n_threads=6,
                 verbose=0,
                 ou)


```

